### 2021-11-13 NSF2021/Description/specificaim1.tex
SA1 with PDF

\subsection{Specific Aim 1: Measuring material properties of amphiphile
  self-assembly}
\label{subsec:specific_aim_1}
Specific Aim 1 proposes a novel mathematical model for particle
self-organization driven by long-range interfacial forces. 
The research aims to provide fundamental insight into the
self-assembly dynamics of amphiphilic particles. These results will
facilitate optimal design of smart materials by tuning the geometry and
properties of the amphiphilic particles.

@@

\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.9in}m{1.9in}m{1.9in}}            
      (a)
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{$t = t_{\text{final}}$}
  \end{tabular}
    \end{center}
  \caption{\label{fig:self-assembly}
    Using the same initial configuration,
    particles self-assemble into distinct morphologies based
    on the boundart condition $g.$ 
    boundary condition leads to a checkerboard pattern.
    In row (a), self-assembly with an amphiphilic boundary
    condition \eqref{eq:bcs}(a) results a bilayer pattern shielding
    the hydrophobic core (yellow, $u > 0$) from the aqueous phase (blue, $u = 0$).
    Row (b) shows particles with a hydrophobic intensity that is
    greater on one side of the particle (magenta) than the other (yellow).
    In row (c), green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    Particles orient like sides to form a checkerboard pattern.
  }
\end{figure}

@@

The ultimate goal is to use numerical simulations to aid
the syntheses soft matter materials.
Applications in this area include cargo-carrying capacity, 
catalytic activity, or stimuli-responsive shape-change to amplify colloid
utility for a wide range of  spanning environmental remediation to drug
delivery \cite{McBr21, HaBr20}.
Here, mathematical models like the kind we propose are crucial
because the model parameters control particle surface properties independently, 
whereas these properties are often difficult to decouple in experiment
~\cite{Bradley2016,Mallory2017,Bradley2017}.

One challenge is translating phenomenological parameters
into measurable quantities.  For example, Janus particles
with uniform surface properties are costly or difficult to
manufacture.  Thus, we can replace the ideal boundary
conditions \eqref{eq:bcs} with a more realistic one like
\begin{equation}
\label{eq:vonMises}
  g_i(\xx) = \sqrt{
  \frac{\exp( \kappa_i\cos(\theta_i(\xx) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}.
\end{equation}
Here, the term under the square root is the von Mises distribution with
director angle $\theta^0_i$ and Bessel function $I_0$ of order zero~\cite{Fisher_1993}.
The square of this function behaves as a normal distribution with variance
like $1/\kappa_i$. We can replicated surface imperfections
by randomizing the concentration $\kappa_i$ and particle radii $r_i$,
for example. Collaboration with experimental groups will
use numerical methods to predict parameter regimes 
to control particle-particle interactions.
In Specific Aim 3, we discuss how to include ions and surface charge.

@@

\begin{wrapfigure}[28]{l}{1.8in}
  \vspace{-4pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} (a) The double well potential $u_0 =
  0.5$, $u_1 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
  and inexact Newton's methods.}
\end{wrapfigure}
Equation~\eqref{eq:SL} takes the form of a steady-state Allen-Cahn
equation. The Allen-Cahn equation is used in phase field formulations of
immisible fluids, for example.
PI BQ has developed fast integral equation solvers for
the Allen-Cahn equation \cite{kro-qua2011, qua2011}.
PI RR and colaborators launched the
simulation and asymptotic analysis of phase field functionals of
membrane elastic energy, devising functionals for for the squared mean
curvature (Willmore) energy~\cite{0951-7715-18-3-016}, spontaneous
curvature~\cite{Du05} and Gaussian curvature energy~\cite{DuEuler}, as
well as energetic variational approaches to coupling membrane elasticity
to fluids~\cite{QiangDu09}.

The key innovation is that when particles are
well-separated, the energetic cost for filling the interstitial region
with $u_0$-phase water is prohibitive and there is a phase shift from
the $u_0$ to $u_1$ water order---the attraction landscape is more or
less identical to the linear response theory. Conversely, when particles
are proximal, it becomes energetically favorable to coalesce
$u_0$-phases. This introduces a line tension for the phase transition
which is an effect not present in the one-dimensional model.

The double-well formulation can potentially 
explain hydrophobic interaction over the full distance
regime~\cite{Lum1999, Lin2005, Meyer2006, Ducker2016}.
Motivated by capillary wetting, Gerhard
Gompper introduced the double-well model for one-dimensional hydrophobic
attraction \cite{GoHaKo94}.
The results improved on previous work~\cite{ErLjCl89,MaRa76},
but the realistic hydrophobic effect is more
complicated and may involve higher-dimensional geometric features. 
Some of the model parameters directly
%connect to long-range decay lengths and energy cost of creating
%hydrophobic-water interfaces with established ranges of
to well-established experimental values. Other
parameters, like the details of the well structure,
are less well-understood and must be fit to the measured data.
This improved model will allow us to more accurately 
simulate colloidal suspensions.
%The experimental data on hydrophobic attraction between 
%is well-established and the literature contains data for a number of
%solvents and hydrocarbon species.
%The experimental measurements
%are well-fit by a double exponential curve in the 10 to 100 nanometer range
%of separation. Below 10~nm, however, there is an amplification of force
%and the attraction follows a much steeper force law not accounted
%for by past phenomenoloical theories~\cite{Lin2005}.

@@

\todo[inline]{Below needs citation and more solid explanation}
Other approaches will be required if the modified phenomenological
theory does not recreate the measured force curves.
One such approach is to consider Gaussian fluctuation. Here, we expand
the free energy
functional around the saddle point solution. The fluctuation energy
then involves calculation of the eigenfunctions of the second variation
of the free energy. This accounts for the fact that experimental
measurements are carried out at finite temperatures, where thermal
fluctuations should ideally also be considered.

>>>>

\subsection{Specific Aim 1: Measuring material properties of amphiphile
  self-assembly}
\label{subsec:specific_aim_1}
Specific Aim 1 proposes a novel mathematical model for particle
self-organization driven by long-range interfacial forces. 
The research aims to provide fundamental insight into the
self-assembly dynamics of amphiphilic particles. These results will
facilitate optimal design of smart materials by tuning the geometry and
properties of the amphiphilic particles.
Applications in this area include cargo-carrying capacity, 
catalytic activity, or stimuli-responsive shape-changes
\cite{McBr21, HaBr20}.

@@

\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.9in}m{1.9in}m{1.9in}}            
      (a)
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{$t = t_{\text{final}}$}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:self-assembly}
    Using the same initial configuration,
    particles self-assemble into distinct morphologies based
    on the boundary condition $g.$ 
    In row (a), self-assembly with an amphiphilic boundary
    condition \eqref{eq:bcs}(a) results a bilayer pattern shielding
    the hydrophobic core (yellow, $u > 0$) from the aqueous phase (blue, $u = 0$).
    Row (b) shows particles with a hydrophobic intensity that is
    greater on one side of the particle (magenta) than the other (yellow).
    In row (c), green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    Particles orient like sides to form a checkerboard pattern.\\
  }
\end{figure}

@@

\begin{wrapfigure}[14]{l}{2in}
  \vspace{-4pt}
  \centering
  \includegraphics[width=2in]{figures/SpecificAim1/OrderPDFs.jpg} 
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:OrderPDFs}
    Probability density functions for
    relative orientation of nearby particles
    for the data in 
    Figure~\ref{fig:self-assembly}, $t = t_{\text{final}}$.
  }
\end{wrapfigure}
The ultimate goal is to use numerical simulations to aid
the syntheses of soft matter materials.
Here, mathematical models like the kind we propose are crucial
because the model parameters control particle surface properties independently, 
whereas these properties are often difficult to decouple in experiment
~\cite{Bradley2016,Mallory2017,Bradley2017}.
The ability to translate phenomenological parameters
into measurable quantities is important. 
For example, Figure~\ref{fig:OrderPDFs} plots
the probability density functions for
relative orientation of nearby particles.
The single peak and broad tails for the blue curve indicates a bilayer structure.
The bimodal red curve indicates multilamellar monolayers. Finally,
The yellow distribution has peaks at $\pm 45^{\circ}$ 
indicating a checkboard pattern. 
As another example, Janus particles
with uniform surface properties are costly or difficult to
manufacture.  Thus, we can replace the ideal boundary
conditions \eqref{eq:bcs} with a more realistic one like
\begin{equation}
\label{eq:vonMises}
  g_i(\xx) = \sqrt{
  \frac{\exp( \kappa_i\cos(\theta_i(\xx) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}.
\end{equation}
Here, the term under the square root is the von Mises distribution with
director angle $\theta^0_i$ and Bessel function $I_0$ of order zero~\cite{Fisher_1993}.
The square of this function behaves as a normal distribution with variance
like $1/\kappa_i$. We can replicated surface imperfections
by randomizing the concentration $\kappa_i$ and particle radii $r_i$,
for example. Collaboration with experimental groups will
use numerical methods to predict parameter regimes 
to control particle-particle interactions.
In Specific Aim 3, we discuss how to include ions and surface charge.

@@

\begin{wrapfigure}[29]{l}{1.8in}
  \vspace{-4pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} (a) The double well potential $u_0 =
  0.5$, $u_1 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
  and inexact Newton's methods.}
\end{wrapfigure}
Equation~\eqref{eq:SL} takes the form of a steady-state Allen-Cahn
equation. The Allen-Cahn equation is used in phase field formulations of
immisible fluids, for example.
PI BQ has developed fast integral equation solvers for
the Allen-Cahn equation \cite{kro-qua2011, qua2011}.
PI RR and colaborators launched the
simulation and asymptotic analysis of phase field functionals of
membrane elastic energy, devising functionals for for the squared mean
curvature (Willmore) energy~\cite{0951-7715-18-3-016}, spontaneous
curvature~\cite{Du05} and Gaussian curvature energy~\cite{DuEuler}, as
well as energetic variational approaches to coupling membrane elasticity
to fluids~\cite{QiangDu09}.

The double-well formulation can potentially 
explain hydrophobic interaction over the full distance
regime~\cite{Lum1999, Lin2005, Meyer2006, Ducker2016}.
Motivated by capillary wetting, Gerhard
Gompper introduced the double-well model for one-dimensional
hydrophobic attraction \cite{GoHaKo94}.
The results improved on previous work~\cite{ErLjCl89,MaRa76},
but a realistic model of hydrophobic effect 
probably involves particle shapes in addition to distances of separation.
The key innovation in the model we propose is that when particles are
well-separated, the energetic cost for filling the interstitial region
with $u_0$-phase water is prohibitive and there is a phase shift from
the $u_0$ to $u_1$ water order---the attraction landscape is more or
less identical to the linear response theory. Conversely, when particles
are proximal, it becomes energetically favorable to coalesce
$u_0$-phases. This introduces a line tension for the phase transition
which is an effect not present in the one-dimensional model.

Some of the model parameters connect directly
%connect to long-range decay lengths and energy cost of creating
%hydrophobic-water interfaces with established ranges of
to well-established experimental values. Other
parameters, like the details of the well structure,
are less well-understood and must be fit to the measured data.
This improved model will allow us to more accurately 
simulate colloidal suspensions.
%The experimental data on hydrophobic attraction between 
%is well-established and the literature contains data for a number of
%solvents and hydrocarbon species.
%The experimental measurements
%are well-fit by a double exponential curve in the 10 to 100 nanometer range
%of separation. Below 10~nm, however, there is an amplification of force
%and the attraction follows a much steeper force law not accounted
%for by past phenomenoloical theories~\cite{Lin2005}.

@@

Other approaches may be required if the modified phenomenological
theory does not recreate measurements.
One such approach is to consider Gaussian fluctuation. Here, we expand
the free energy
functional around the saddle point solution. The fluctuation energy
then involves calculation of the eigenfunctions of the second variation
of the free energy. This accounts for the fact that experimental
measurements are carried out at finite temperatures, where thermal
fluctuations should ideally also be considered. 
\todo[inline]{Above, needs citation and more solid explanation}
